---
title: "BASEBALL"
AUTHOR: Alfred Cheung

execute: 
  message: false
  echo: false

---

```{r}
#| echo: FALSE
#| message: FALSE
library(tidyverse,ggplot2)
```


```{r}
#| warning: FALSE


df = read.csv("retro_weather2.csv")

df = df %>%
  # Make >95 the highest bin  **Adding Why for the people who didnt know how to code**
  mutate(wtemp_avg_bin12=ifelse(wtemp_avg_bin13==1,1,wtemp_avg_bin12)) %>%
  # create bins
  mutate(temp_bin = case_when(wtemp_avg_bin1==1 ~ '<45°',
                              wtemp_avg_bin2==1 ~ '45-50°',
                              wtemp_avg_bin3==1 ~ '50-55°',
                              wtemp_avg_bin4==1 ~ '55-60°',
                              wtemp_avg_bin5==1 ~ '60-65°',
                              wtemp_avg_bin6==1 ~ '65-70°',
                              wtemp_avg_bin7==1 ~ '70-75°',
                              wtemp_avg_bin8==1 ~ '75-80°',
                              wtemp_avg_bin9==1 ~ '80-85°',
                              wtemp_avg_bin10==1 ~ '85-90°',
                              wtemp_avg_bin11==1 ~ '90-95°',
                              wtemp_avg_bin12==1 ~ '>95°')) %>%
  # filter out NAs introduced by having Canada in the data
  filter(!is.na(wtemp_avg)) %>%
  # group by bin and count number of obs
  count(temp_bin) 

# set the bins in a character vector
bins = c('<45°', '45-50°', '50-55°', '55-60°', '60-65°', '65-70°',
         '70-75°', '75-80°', '80-85°', '85-90°', '90-95°', '>95°')

# create factor levels for the temp bins
df$temp_bin = factor(df$temp_bin, levels = bins)

# plot the histogram ------------------------------------------------------
gg = ggplot(df, aes(temp_bin, n)) +
  geom_col(width = .85, fill = "#005A9C")  + 
  scale_y_continuous(breaks = c(0, 5000, 10000, 15000, 20000),
                     labels = c('0', '5,000', '10,000', '15,000', 
                                '20,000'),
                     name = 'Count', 
                     expand = c(0, 0)) +
  scale_x_discrete(name = 'Temperature Bin',
                   expand = c(0, 0)) +
  theme_minimal() +
  coord_cartesian(clip = "off") +
  theme(panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = 'grey'),
        axis.ticks.x = element_line())

# display

# presentation version
presentation = gg +
  labs(title = 'Number of MLB games in each 5-degree temperature bin, 1950-2000') +
  theme(axis.text = element_text(family = 'Lato', size = 12),
        axis.title = element_text(family = 'Lato', size = 12),
        plot.title = element_text(lineheight = 1.1, family = 'Lato', size = 20),
        plot.title.position = 'plot')

#display
presentation
```

This chart displays the distribution of MLB games across different temperature ranges from 1950 to 2021, revealing a clear preference for moderate to warm playing conditions. The data shows that baseball games are most frequently played in temperatures between 65-80°F, with the peak occurring in the 70-75°F range at over 22,000 games. Game frequency drops significantly at temperature extremes, with very few games played below 45°F or above 95°F, reflecting both player comfort and fan attendance considerations.



```{r}
# Description -------------------------------------------------------------

# This script plots the average attendance data as a share of stadium capacity across the sample period

# -------------------------------------------------------------------------

# install/load other packages
library(tidyverse,
       viridis
       )

# turn off scientific notation
options(scipen = 999)

# load the retrosheet data
retro = read_csv("retro_weather2.csv")


# transform the data ------------------------------------------------------
retro = retro %>%
  # create a day-of-the-year variable and make it an R date
  mutate(week = week(date),
         month = month(date),
         month_day = format(date, "%B %d"),
         month_day = as.Date(month_day, "%B %d"),
         capacity_share = attendance / capacity) %>%
  # group by month_day-team
  group_by(month, year) %>%
  # summarise for average attendance
  summarise_all(funs(mean), na.rm=TRUE)

# plot the data -----------------------------------------------------------
ggplot(retro, aes(date, capacity_share)) +
  geom_line(color = '#3D8361') +  
  scale_y_continuous(
    name = "Ballpark Capacity",
    breaks = c(.1, .2, .3, .4, .5, .6, .7, .8, .9, 1),
    labels = scales::percent_format(accuracy = 1)) + 
  scale_x_date(breaks = seq(as.Date("1950-01-01"), as.Date("2021-01-01"),
                            by = "5 years"),
               date_labels = '%Y') +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        panel.grid.major = element_line())
  

```